# Random landscape generation

Trying to generate random landscapes from homogeneous gradient to very patchy, 
as well as have a metric to describe the heterogenaity.

```{r}
library(RandomFields) # generate Gaussian random fields
library(geoR) # matern covariance function
library(spdep) # Moran's I
library(fractaldim) # fractal dimension

```

## Examples

First try the examples to generate gaussian random fields

All models
```{r}
PrintModelList()
```


symmetric stable model, also called "powered exponential model"
```{r}
model <- "stable"
mean <- 0
variance <- 4
nugget <- 1
scale <- 10
alpha <- 1 ## see help("CovarianceFct") for additional
## parameters of the covariance functions
step <- 1 ## nicer, but also time consuming if step <- 0.1
x <- seq(0, 50, step)
y <- seq(0, 50, step)
f <- GaussRF(x=x, y=y, model=model, grid=TRUE,
param=c(mean, variance, nugget, scale, alpha))
image(x, y, f)
```

Brownian motion (using Stein’s method)
```{r}
step <- 0.3 ## nicer, but also time consuming if step = 0.1
x <- seq(0, 20, step)
kappa <- 1 # in [0,2)
z <- GaussRF(x=x, y=x, grid=TRUE, model=list("fractalB", kappa))
image(z,zlim=c(-3,3))
```


Matern
```{r}
step = 1
x = seq(0, 50, step)
z = GaussRF(x = x, y = x, grid = TRUE, model = "matern", param = c(mean=0, variance=1, nugget=0, scale=1, nu=0.1))
image(z)
```

Now try manipulate for different parameters
```{r}
library(manipulate)
library(RandomFields) # generate Gaussian random fields
library(fields) # image.plot
habitatCols = c(colorRampPalette(c("gainsboro", "darkolivegreen3"))(10), colorRampPalette(c("darkolivegreen3", "darkolivegreen"))(64))

plotMatern = function(mean, scale, nu)
{
	x = seq(0, 50, 1)
	z = GaussRF(x = x, y = x, grid = TRUE, model = "matern", param = c(mean=mean, variance=1, nugget=0, scale=scale, nu=nu))
	z = normalizeZ(z)
	print(range(z))
	
	par(mar = c(5, 1, 1, 1))
	image(z, col = habitatCols)
	image.plot(legend.only = TRUE, zlim = range(z), horizontal = TRUE, col = habitatCols)
}

normalizeZ = function(z, total = 1)
{
	# make positive
	z[z < 0] = 0
	
	if (sum(z) == 0) print("Invalid field, all 0")

	# sum to one (or other amount for larger landscapes)
	z = z * total / sum(z) 
	
	return(z)
}

manipulate(plotMatern(mean, scale, nu),
		   mean = slider(-2.0, 2.0),
		   scale = slider(0.001, 100),
		   nu = slider(0.001, 200))
```

Try model fitting the results for different means (i.e. different amounts cut off to 0)
```{r,  cache=TRUE}
	x = seq(0, 50, 1)
	z = GaussRF(x = x, y = x, grid = TRUE, model = "matern", param = c(mean=2, variance=1, nugget=0, scale=5, nu=1.5))
	z = normalizeZ(z)
#	v = fitvario(as.matrix(expand.grid(x, x)), data=z, model="matern", param = c(NA, NA, 0, NA, NA), 
#				 lower = c(0.1, 0.1, 0.1), upper = c(2, 10, 2))
#	summary(v)
```


Now generate a bunch and calculate some statistics
```{r}
generateLandscape = function(mean, scale, dim = 50, totalSum = 1)
{
	x = seq(1, dim, 1) # should be 1 to 50, eg
	z = NULL
	
	tries = 20
	i = 1
	while (i < tries)
	{
		# matern - nu parameter not interesting
		# exponential, when nu = 1/2
		# gauss (aka squared exponential), when nu = inf
		z = GaussRF(x = x, y = x, grid = TRUE, model = "exponential", param = c(mean=mean, variance=1, nugget=0, scale=scale))
		if (any(z > 0))
		{
			break	
		}
		i = i + 1 
	}

	if (is.null(z))
	{
		# TODO, eventually warn and fail gracefully? with uniform landscape
		stop("z all negative after", tries, "tries")
	}
	z = normalizeZ(z, totalSum)	
	return(z)
}

plotLandscape = function(z, mean, scale)
{
	image(z, col = habitatCols)
	title( paste("m =", mean, "s =", scale))
	text(x = 0, y = 49, labels = range(z), pos = 4)
}

calculateNeighbors = function(x)
{
	# for Moran's I since these depend on the size of the grid, not the data
	coords = as.matrix(expand.grid(x, x))
	neigh.nb = dnearneigh(coords, 0, 2 * max(x), longlat=FALSE) # warning, this is really slow
	neigh.dist = nbdists(neigh.nb, coords, longlat=FALSE)
	inverse = lapply(neigh.dist, function(x) (1/x))
	class(inverse) = c("nbdist") # class gets lost with lapply, needed?
	neigh.listw = nb2listw(neigh.nb, glist = inverse, style="W", zero.policy=TRUE) # also slow
	# what style? 
	#W is row standardised (sums over all links to n), 
	#C is globally standardised (sums over all links to n), 
	#U is equal to C divided by the number of neighbours (sums over all links to unity)
	szero = Szero(neigh.listw)
	
	return(list(nb = neigh.nb, listw = neigh.listw, szero = szero))
}

calculateMetrics = function(paramData, nReplicates, neighbors = NULL, plot = TRUE)
{
	if (is.null(neighbors))
	{
		x = seq( from = 1, to = 50, by = 1) # start from 0 or 1?
		neighbors = calculateNeighbors(x)
	}
	
	paramData = cbind(paramData, sdMean = NA, sdU = NA, sdL = NA,  moransIMean = NA, moransIU = NA, moransIL = NA, fractalMean = NA, fractalU = NA, fractalL = NA)
	
	rows = 6
	cols = 4
	
	for (i in 1:nrow(paramData))
	{
		land = lapply(1:nReplicates, function(x) generateLandscape(mean = paramData[i, "mean"], 
																 scale = paramData[i, "scale"]) )
		
		if (plot)
		{
			if ((i %% (rows * cols)) == 1) 
			{
				pdf(file = paste("landParams", (i - 1) / (rows * cols) + 1, ".pdf", sep = ""), width = 8, height = 12)
				par(mfrow = c(rows, cols), mar = c(1, 1, 3, 1))
			}
			
			plotLandscape(land[[1]], paramData[i, "mean"], paramData[i, "scale"])
			
			if ((i %% (rows * cols)) == 0)
			{
				dev.off()
			}
		}
		sd = sapply(land, function(x) sd(as.vector(x)))
		moransI = unlist( sapply(land, function(x) moran(x, neighbors$listw, length(neighbors$nb), neighbors$szero))[1,] )
		fractal = sapply(land, function(x) fd.estim.isotropic(x, p.index = 1, direction='hv', plot.loglog = FALSE)$fd)
		
		paramData[i, 3:11] = c(mean(sd), quantile(sd, probs = c(0.975, 0.025)), 
							   mean(moransI), quantile(moransI, probs = c(0.975, 0.025)), 
							   mean(fractal), quantile(fractal, probs = c(0.975, 0.025)))
	}
	return(paramData)
}

evalLandscapeParameters = function(neighbors = NULL)
{
	mean = c(-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5)
	scale = c(0.5, 1:10, 15, 20, 25, 50)
	
	landResults = expand.grid(mean = mean, scale = scale)
	landResults = calculateMetrics(landResults, 20, neighbors)
	return(landResults)
}

x = seq( from = 1, to = 50, by = 1) # start from 0 or 1?
neighbors = calculateNeighbors(x)
lands = evalLandscapeParameters(neighbors)

plotConf = function(data, colIdx, ylab)
{
	require(plotrix)
	require(RColorBrewer)
	cols = brewer.pal(10, "BrBG")
	meanN = as.numeric(factor(data$mean))
	scaleN = as.numeric(factor(data$scale))
	upperCI = data[,colIdx + 1]
	lowerCI = data[,colIdx + 2]
	xdiff = rep(seq(from = -0.4, to = 0.4, length = length(unique(meanN))), length(unique(scaleN)))
	
	plotCI(x = scaleN + xdiff, y = data[,colIdx], ui = upperCI, li = lowerCI, 
		   col = cols[meanN], sfrac = 0, pch = 16, xlab = "scale", ylab = ylab, xaxt = "n")
	axis(side = 1, at = scaleN, labels = factor(data$scale))
	
	legend("topleft", legend = levels(factor(data$mean)), fill = cols[1:length(unique(meanN))], 
		   ncol = 3, bty = "n")
}

plotConf(lands, 3, "Std Dev")
plotConf(lands, 6, "Moran's I")
plotConf(lands, 9, "Fractal Dim")
```

Now to generate some landscapes
```{r}
generateLandscapes = function(parameters, neighbors = NULL)
{
	if (is.null(neighbors))
	{
		x = seq( from = 1, to = 50, by = 1) 
		neighbors = calculateNeighbors(x)
	}

	parameters = cbind(parameters, sd = NA, moransI = NA, fractal = NA)
	
	for (i in 1:nrow(parameters))
	{
		land = generateLandscape(mean = parameters[i, "mean"],
								 scale = parameters[i, "scale"])
		parameters[i, "sd"] = sd(as.vector(land))
		parameters[i, "moransI"] = moran(land, neighbors$listw, length(neighbors$nb), neighbors$szero)$I
		parameters[i, "fractal"] = fd.estim.isotropic(land, p.index = 1, direction='hv', plot.loglog = FALSE)$fd

		write.csv(land, file = paste("land/land", parameters[i, "id"], ".csv", sep = ""), quote = FALSE, row.names = FALSE)
	}
	write.csv(parameters, file = "landParameters.csv", quote = FALSE, row.names = FALSE)
	return(parameters)
}

scales = rep(c(1, 1, 5, 5, 5, 20), each = 10 * 5)
nus = rep(c(1, 100, 1, 2, 100, 2), each = 10 * 5)
means = rep( c(-1, -0.5, 0, 0.5, 1), times = 6 * 10)
ids = 1:length(scales)
params = data.frame(scale = scales, nu = nus, mean = means, id = ids)

#alternatively
params = expand.grid(mean = c(-2, -1.5, -1, -0.5, 0, 0.5, 1, 1.5), scale = c(0.5, 1:10, 15, 20, 25, 50))
params = params[rep(1:nrow(params), times = 20),]
params$id = 1:nrow(params)

params = generateLandscapes(params, neighbors)

# create list for properties file
names = paste("land", 1:2400, ".csv", sep = "")
cat(names, sep = ";" )
```

# landscape for introduced predators
```{r}
background = generateLandscape(mean = 0, scale = 10, dim = 100) * 4
better = GaussRF(x = seq(51, 75, 1), y = seq(26, 75, 1), grid = TRUE, 
				 model = "exponential", param = c(mean=-1.5, variance=1, nugget=0, scale=10))
better = normalizeZ(better)
layer = matrix(0, nrow = 100, ncol = 100)
layer[seq(51, 75, 1), seq(26, 75, 1)] = better
landscape = background + layer
plotLandscape(landscape, mean = "var", scale = 10)
write.csv(landscape, file = "model/land/landPredInt.csv", quote = FALSE, row.names = FALSE)

par(mfrow=c(2,2))
plotLandscape(as.matrix(read.csv("model/land/landPred.csv", header = TRUE)), mean = "var", scale = "var")
plotLandscape(as.matrix(read.csv("model/land/landPred2.csv", header = TRUE)), mean = "var", scale = "var")
plotLandscape(as.matrix(read.csv("model/land/landPredInt.csv", header = TRUE)), mean = "var", scale = "var")
plotLandscape(t(as.matrix(read.csv("model/land/yellJul2005.csv", header = TRUE)))/1500, mean = "var", scale = "var")


# try exponentiating to increase contrast
plotLandscape(exp(as.matrix(read.csv("model/land/yellJul2005.csv", header = TRUE))*0.02), mean = 0.02, scale = "var")
plotLandscape(exp(as.matrix(read.csv("model/land/yellJul2005.csv", header = TRUE))*0.03), mean = 0.03, scale = "var")
plotLandscape(exp(as.matrix(read.csv("model/land/yellJul2005.csv", header = TRUE))*0.04), mean = 0.04, scale = "var")
plotLandscape(exp(as.matrix(read.csv("model/land/yellJul2005.csv", header = TRUE))*0.05), mean = 0.05, scale = "var")
```


# landscape for naive foragers, 4 levels of patch quality
```{r}
# properties of some existing, mean = -1.5, scale = 2, 5, 10, 50
land18 = read.csv("model/land/land18.csv", header = TRUE)
land42 = read.csv("model/land/land42.csv", header = TRUE)
land82 = read.csv("model/land/land82.csv", header = TRUE)
land114 = read.csv("model/land/land114.csv", header = TRUE)
lands = list(land18, land42, land82, land114)
par(mfrow=c(2,2))
for(i in 1:4)
{
		hist(lands[[i]][lands[[i]]>0], main = sum(lands[[i]] == 0)/nrow(lands[[i]])/ncol(lands[[i]]))
}

layer1 = generateLandscape(mean = -2, scale = 10, dim = 250, totalSum = 10)
layer2 = generateLandscape(mean = -2, scale = 10, dim = 250, totalSum = 7.5)
layer3 = generateLandscape(mean = -1.5, scale = 10, dim = 250, totalSum = 5)
layer4 = generateLandscape(mean = -1.5, scale = 10, dim = 250, totalSum = 2.5)
layer1[100:150, 100:150] = 0
layer2[100:150, 100:150] = 0
landscape = layer1 + layer2 + layer3 + layer4
landscape = normalizeZ(landscape, total = 25)
plotLandscape(landscape)
hist(landscape[landscape>0], main = sum(landscape == 0)/nrow(landscape)/ncol(landscape))
image(landscape, col = c("white", "black"), breaks = c(-1, 0, 10))
write.csv(landscape, file = "model/land/landTransLoc.csv", quote = FALSE, row.names = FALSE)

# bigger landscape just 82 stiched together
landscape82 = NULL
landscape82 = cbind(land82, land82, land82, land82, land82)
landscape82 = rbind(landscape82, landscape82, landscape82, landscape82, landscape82)
write.csv(landscape82, file = "model/land/landTransLoc82.csv", quote = FALSE, row.names = FALSE)

landscape = read.csv("model/land/landTransLoc.csv", header = TRUE)
landscape82test = read.csv("model/land/landTransLoc82.csv", header = TRUE)

```
